---
title: "Assignment 2: Coding Basics"
author: "Jaleesia Amos"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 

# Generate a sequence of numbers from one to 100, increasing by fours.
seq_by_four <- seq(1, 100, 4) 
seq_by_four

#2. 

# Mean and median of seq_by_four
mean(seq_by_four)
median(seq_by_four)

#3. 

# Determine whether the mean is greater than the median
mean(seq_by_four) > median(seq_by_four)

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
# Vector of names of student
student <- c("Jessica", "Jasmine", "John", "Jackie")

# Test scores out of a total 100 points
test_scores <- c(98, 80, 95, 49)

# Whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50
pass <- c("TRUE", "TRUE", "TRUE", "FALSE")

# Combine vectors into data frame with informative titles (vectors given informative titles. no need to change)

student_scores <- data.frame(student, test_scores, pass)
student_scores


```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:  A matrix has a set of rows and columns of the same data type. A data frame is similar to a matrix, but you can have columns of different data types.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}

# Function for whether a given test score is a passing grade of 50 or above (TRUE or FALSE)
# Creation of function
pass_fail <- function(x) {
  ifelse(x>50, print("TRUE"), print("FALSE"))
}

# Test function on vector
pass_fail(test_scores)

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: I used the 'ifelse' statement, but technically both options should work because they do the samething. 
