---
title: "Assignment 5: Data Visualization"
author: "Jaleesia Amos"
date: "Spring 2023"
output: pdf_document
geometry: margin=2.54cm
fig_height: 3
fig_width: 4
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1.  Rename this file `<FirstLast>_A05_DataVisualization.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure your code is tidy; use line breaks to ensure your code fits in the knitted output.
5. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.

---

## Set up your session 

1. Set up your session. Load the tidyverse, lubridate, here & cowplot packages, and verify your home directory. Upload the NTL-LTER processed data files for nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy `NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv` version) and the processed data file for the Niwot Ridge litter dataset (use the `NEON_NIWO_Litter_mass_trap_Processed.csv` version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r initialize project}
#1 

#----Load packages into session & download if not already installed----#
pacman::p_load(tidyverse, lubridate, here, cowplot)

#----Check working directory----#
getwd()

#----Read in dataset 1: Nutrients dataset-----#
PeterPaul_processed <- read.csv("./Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)

#----Read in dataset 2: Litter dataset------#
litter_processed <- read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)

#2 

#----Checking structure of Nutrients for date format------#
str(PeterPaul_processed)

#----Checking structure of Litter for date format------#
str(litter_processed)

#_____Date Conversion: Nutrients dataset______#
PeterPaul_processed$sampledate <- ymd(PeterPaul_processed$sampledate)

#_____Date Conversion: Litter dataset______#
litter_processed$collectDate <- ymd(litter_processed$collectDate)

```
## Define your theme
3. Build a theme and set it as your default theme. Customize the look of at least two of the following: 
 * Plot background
 * Plot title
 * Axis labels
 * Axis ticks/gridlines
 * Legend

```{r build plot theme}
#3
#----Custom theme: change title, axis title, and legend----#
jatheme <- theme_bw(base_size = 12) + 
  theme(plot.title = element_text(face ="bold", hjust = 0.5),
        axis.title = element_text(face = "bold"), 
        legend.position = "right")

#-----Set theme for environment-----#
theme_set(jatheme)

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best practices for data visualization. Ensure your theme, color palettes, axes, and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with separate aesthetics for Peter and Paul lakes. Add a line of best fit and color it black. Adjust your axes to hide extreme values (hint: change the limits using `xlim()` and/or `ylim()`).

```{r plot total P vs PO4}
#4 
#----Couldn't create line of best of fit through geom_line or geom_abline----#
LinMod <- lm(tp_ug~po4, data = PeterPaul_processed)

#----get intercept and slope value
coeff <- coefficients(LinMod)
intercept <- coeff[1]
slope <- coeff[2]

#----Create plot---#
ggplot(data = PeterPaul_processed) +
  geom_point(aes(x = po4, y = tp_ug, color = lakename)) +
#----Change axis title---#
  xlab("Phosphate") +
  ylab("Total Phosphorus") +
#----Add plot title----#  
  ggtitle("Total phosphorus vs phosphate for Paul and Peter Lake") +
#----Adjust axis scale----#
  xlim(0, 50) +
  geom_abline(intercept = intercept, slope = slope, color = "black")

 
```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and (c) TN, with month as the x axis and lake as a color aesthetic. Then, create a cowplot that combines the three graphs. Make sure that only one legend is present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;see <https://r-lang.com/month-abb-in-r-with-example>

```{r Create boxplots}
#5 
#-----5a. boxplot of temperature------#
tempPlot <- ggplot(data = PeterPaul_processed)  +
  geom_boxplot(aes(x = as.factor(month), y = temperature_C, 
                   color = lakename)) 
tempPlot

#-----5b. boxplot of TP------#
tpPlot <- ggplot(data = PeterPaul_processed) +
  geom_boxplot(aes(x = as.factor(month), y = tp_ug, 
                   color = lakename)) 
tpPlot

#-----5c. boxplot of TN------#
tnPlot <- ggplot(data = PeterPaul_processed) +
  geom_boxplot(aes(x = as.factor(month), y = tn_ug, 
                   color = lakename)) 
tnPlot

#------Remove legends from individual plots-----#
prow <- plot_grid(tempPlot + theme(legend.position = "none"),
                  tpPlot + theme(legend.position = "none"),
                  tnPlot + theme(legend.position = "none"),
                  nrow = 3,
                  axis = "b",
                  align = "v")

#------Get legend from one plot------#
legend <- get_legend(tempPlot)

#------Combine plots into one plot-----#
plot_grid(prow, legend)
```

Question: What do you observe about the variables of interest over seasons and between lakes? 

> Answer: There is a positive correlation between the temperature and the concentration levels of tp and tn. During the summer when temperatures are higher, tp and tn concentrations are also higher. Between Paul and Peter Lake, the tp and tn concentrations are higher in Peter lake. The temperatures between Paul and Peter Lake are nearly the same.

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the "Needles" functional group. Plot the dry mass of needle litter by date and separate by NLCD class with a color aesthetic. (no need to adjust the name of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into three facets rather than separated by color. 

```{r Plot litter}
#6
#-----Plot by functional group-----#
ggplot(data = litter_processed %>%
         filter(functionalGroup == "Needles")) +
         geom_point(aes(x = collectDate, y = dryMass, color = nlcdClass, fill = nlcdClass))

#7
#-----Create same plot with facets-----#
ggplot(data = litter_processed %>%
         filter(functionalGroup == "Needles")) +
         geom_point(aes(x = collectDate, y = dryMass, 
                      color = collectDate)) + 
                      facet_wrap(vars(nlcdClass))

```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: Plot 7 is more effective in this case, because the facet feature allows us to more clearly distinguish between the dry mass of needles by nlcdClass. For example, it is easier to determine that shrub scrub is produces the least amount needles across all years that data were collected.